---
title: "Section 3 Video 2 : Implementing and Visualizing PCA"
author: "Antoine Pissoort"
output: 
  html_document:  
    theme: readable  # Document Style
    highlight: tango # Syntax Highlighting
    number_sections: yes # Table Of Content numbering
    toc: yes  # Table Of Content
    toc_depth: 3  # Table Of Content depth
    df_print: paged # Print the data frames in a html table
---
  
  
# Automatic Packages Installing and Loading 
  
```{r setup}
knitr::opts_chunk$set(echo       = TRUE,
                      fig.height = 3.5,
                      fig.width  = 6)

packages <- c("tidyverse", "FactoMineR", "factoextra")

## Install packages From CRAN if they are not installed.
for (p in packages) {
  if (p %in% rownames(installed.packages()) == FALSE) {
    install.packages(p, repos = 'http://cran.us.r-project.org')
  }
}
## Load the packages
for (p in packages)
  suppressPackageStartupMessages(
    library(p, quietly = TRUE, character.only = TRUE ) 
  )
```


# Load Decathlon Dataset  

```{r}
data(decathlon2, package = "factoextra")

# Take 10 events as features. Remove Rank, Points and Competition
(decathlon2_features <- decathlon2[, 1:10])
```


# Run PCA

```{r}
pca_decathlon <- PCA(
  X          = decathlon2_features,
  graph      = FALSE,
  scale.unit = TRUE,
  ncp        = 4
)
# Print the structure of the PCA object
pca_decathlon
```


# Eigenvalues: what percentage of variance is explained by the PC's ?  

```{r}
fviz_eig(
  X         = pca_decathlon,
  addlabels = TRUE,
  ylim      = c(0, 4),
  ncp       = 4,
  choice    = "eigenvalue"
)
```

The **percentage of variance explained** $\tau_m$ by the m first PC's is $$\tau_m = \frac{1}{p}(\lambda_1+\ldots+\lambda_m) \times 100$$

```{r}
fviz_eig(
  X         = pca_decathlon,
  addlabels = TRUE,
  ylim      = c(0, 40),
  ncp       = 4,
  choice    = "variance"
)
```

**Cumulative**
```{r}
pca_eig4_df <- data.frame(pca_decathlon$eig) %>% 
  rownames_to_column(var = "component") %>% 
  slice(1:4)

ggplot(pca_eig4_df, aes(x = component)) +
  geom_bar(aes(y = percentage.of.variance),
           stat = "identity") +
  geom_point(aes(y = cumulative.percentage.of.variance)) +
  geom_path(aes(y = cumulative.percentage.of.variance, group = 1))
```



# Visualization of the Features in 2D

* `coord`   : coordinates on the dimensions
* `cor`     : correlations with the dimensions
* `cos2`    : quality of representation on the dimensions
* `contrib` : Contributions to the dimensions

```{r}
pca_decathlon$var
```


## Quality of representation : cos2

```{r}
fviz_cos2(pca_decathlon, choice = "var", axes = 1:2)
```


## Correlation circle 

```{r}
fviz_pca_var(pca_decathlon, repel = TRUE)
```

```{r}
# Same but coloured by cos2 values
fviz_pca_var(
  X             = pca_decathlon,
  col.var       = "cos2",
  gradient.cols = c("orange","red", "blue"),
  repel         = TRUE 
)
```

## Contributions of features to principal components

* **1st component**
```{r}
fviz_contrib(pca_decathlon, choice = "var", axes = 1, top = 10)
```

* **2n component**
```{r}
fviz_contrib(pca_decathlon, choice = "var", axes = 2, top = 10)
```

**Correlation circle coulour by contributions**
```{r}
fviz_pca_var(
  X             = pca_decathlon,
  col.var       = "contrib",
  gradient.cols = c("orange", "red", "blue"),
  repel         = TRUE
)
```




# Visualization of the Individuals in 2D

* `coord`   : coordinates on the dimensions
* `cos2`    : quality of representation on the dimensions
* `contrib` : Contributions to the dimensions

```{r}
str(pca_decathlon$ind)
```


```{r}
fviz_pca_ind(
  X             = pca_decathlon,
  col.ind       = "cos2",
  pointsize     = "contrib",
  gradient.cols = c("orange", "red", "blue"),
  repel         = TRUE
) + theme(
  plot.title = element_text(
    size   =  15,
    colour = "#33666C",
    face   = "bold",
    hjust  = 0.5
  ),
  axis.title = element_text(size = 11, colour = "#33666C", face = "bold"),
  legend.background = element_rect(colour = "black")
)
```


# Visualization of individuals and features in 2D : biplot

```{r}
fviz_pca_biplot(
  X             = pca_decathlon,
  repel         = TRUE,
  col.var       = "black",
  col.ind       = "cos2", 
  gradient.cols = c("orange", "red", "blue")
)
```


```{r}
decathlon2 %>% rownames_to_column %>% arrange(desc(High.jump))
```
