---
title: "Section 4 Video 2: Computing Item Sets : association rules"
author: "Antoine Pissoort"
output: 
  html_document:  
    theme: readable  # Document Style
    highlight: tango # Syntax Highlighting
    number_sections: yes # Table Of Content numbering
    toc: yes  # Table Of Content
    toc_depth: 3  # Table Of Content depth
    df_print: paged # Print the data frames in a html table
    code_folding: hide  # Hide the code in the report  
---


# Automatic Packages Installing and Loading 
  
```{r setup}
knitr::opts_chunk$set(echo       = TRUE,
                      fig.height = 3.5,
                      fig.width  = 6)

packages <- c("tidyverse", "plyr", "arules")

## Install packages From CRAN if they are not installed.
for (p in packages) {
  if (p %in% rownames(installed.packages()) == FALSE) {
    install.packages(p, repos = 'http://cran.us.r-project.org')
  }
}
## Load the packages
for (p in packages)
  suppressPackageStartupMessages(
    library(p, quietly = TRUE, character.only = TRUE ) 
  )
```


# Read Data

```{r}
(data_online_final <- readRDS(file = "data_online_final.RDS")) 
```


# Compute the transaction dataframe 
 
```{r}
(item_data <- ddply(data_online_final, c("CustomerID", "InvoiceDate"),
                  function(df) paste(df$Description, collapse = ",")))
```

```{r, results = FALSE}
item_data <-  item_data %>% select(-CustomerID, -InvoiceDate)
colnames(item_data) <- c("items")
item_data
```


# Write to csv file 

```{r, eval = FALSE}
write.csv(item_data, "market_basket.csv", quote = FALSE, row.names = TRUE)
```
 

# Read as a transaction object 

```{r}
item_data_tr <- read.transactions(
  file          = 'market_basket.csv',
  format        = 'basket',
  sep           = ',',
  quote         = "", 
  rm.duplicates = TRUE,
  skip          = 1
)
```

```{r}
summary(item_data_tr)
```

Compute the number of items that were purchased : 
`r 18441 * 22347 * 0.0009914608`


## Inspect some item sets

```{r}
item_data_tr %>% head(n = 5) %>% inspect
```


# Apriori algorithm 

1. Identify all item sets that meet a minimum support threshold
2. Create rules from these items that meet a minimum confidence threshold
 
```{r}
rules_apriori <- item_data_tr %>%
  apriori(parameter = list(supp = 0.005, conf = 0.8)) %>%
  sort(by = 'confidence')
```


## Summary of the rules 

```{r}
summary(rules_apriori)
```


## Rules with most confidence 

```{r}
rules_apriori %>% head(n = 5) %>% inspect
```

* 100% customers who bought GARAGE DESIGN also bought KEY FOB.
* 100% customers who bought ELEPHANT also bought BIRTHDAY CARD. 
* ...


## Rules with least confidence 

```{r}
rules_apriori %>% tail(n = 5) %>% inspect
```


## Rules sorted by lift 

$$Lift(X\rightarrow Y) = \frac{Confidence(X \rightarrow Y)}{Support(Y)}$$

```{r}
rules_apriori %>% sort(by = "lift") %>%  head(n = 5) %>% inspect
```

