---
title: "Section 4 Video 3: Visualizing Items Sets"
author: "Antoine Pissoort"
output: 
  html_document:  
    theme: readable  # Document Style
    highlight: tango # Syntax Highlighting
    number_sections: yes # Table Of Content numbering
    toc: yes  # Table Of Content
    toc_depth: 3  # Table Of Content depth
    df_print: paged # Print the data frames in a html table
    code_folding: hide  # Hide the code in the report  
---
  

# Automatic Packages Installing and Loading 
  
```{r setup}
knitr::opts_chunk$set(echo       = TRUE,
                      fig.height = 3.5,
                      fig.width  = 6)

packages <- c("tidyverse", 'arules', "arulesViz")

## Install packages From CRAN if they are not installed.
for (p in packages) {
  if (p %in% rownames(installed.packages()) == FALSE) {
    install.packages(p, repos = 'http://cran.us.r-project.org')
  }
}
## Load the packages
for (p in packages)
  suppressPackageStartupMessages(
    library(p, quietly = TRUE, character.only = TRUE ) 
  )
```


# Read the transaction object 

```{r}
item_data_tr <- read.transactions(
  file          = 'market_basket.csv',
  format        = 'basket',
  sep           = ',',
  quote         = "", 
  rm.duplicates = TRUE,
  skip          = 1
)
```


## Summary 

```{r}
summary(item_data_tr)
```


## Inspect 

```{r}
set.seed(1234)
item_data_tr[sample(1:length(item_data_tr), 3)] %>% inspect
```


# Check the item frequency plot

```{r, fig.height=5}
itemFrequencyPlot(x     = item_data_tr,
                  topN  = 12,
                  # support =
                  type  = 'absolute',
                  horiz = TRUE)
```


# Apriori algorithm 

```{r}
rules_apriori <- item_data_tr %>%
  apriori(parameter = list(supp = 0.005, conf = 0.8)) %>%
  sort(by = c('confidence', "lift"))
```


# Inspect the top 10 rules 

```{r}
inspect(rules_apriori[1:10])
```

# Plot the top 10 rules 

```{r}
plot(rules_apriori[1:10])
```

# Plot the top 100 rules 

```{r}
plot(rules_apriori[1:100])
```

# Plot all the rules 

```{r}
plot(rules_apriori)
```

# Plot method of the arulesViz package

```{r, eval = FALSE}
plot(
  x,
  method      = NULL,
  measure     = "support",
  shading     = "lift",
  control     = NULL,
  engine      = "default"
)
```

* **method** :  "scatterplot", "two-key plot", "matrix", "matrix3D", "mosaic", "doubledecker", "graph", "paracoord" or "grouped", "iplots
* **measure** :  "support", "confidence", "lift", "order"
* **shading** : "support", "confidence", "lift"
* **control** : a list of control parameters for the plot depending on the used visualization method
* **engine**  : plotting engine used to render the plot. Can be "interactive", "3d", ... 


## Plot of the lift vs Support coloured by confidence 

```{r}
plot(rules_apriori,
     measure = c("support", "lift"),
     shading = "confidence")
```


## Plot a graph for the top 10 rules 

```{r}
plot(rules_apriori[1:10], method = "graph")
```

## Plot a graph for the top 10 rules 

```{r, fig.width=7, fig.height=5}
plot(rules_apriori[1:10], method = "grouped")
```


## Plot a 3D matrix for all the rules

```{r, results = "hide"}
plot(rules_apriori, method = "matrix", engine = "3d", measure = "lift")
```

## Plot parallel coordinates for top 10 rules

```{r}
plot(rules_apriori[1:10],
     method  = "paracoord",
     control = list(reorder = TRUE))

```



## Dynamic "two-key" plot for all the rules 

```{r}
plotly_arules(rules_apriori, method = "two-key plot")
```




# Simplify the analysis with Text Mining

Video 2 Section 2. 

