---
title: "Section 2 Video 4 : Training Your Model"
author: "Antoine Pissoort"
output: 
  html_document:  
    theme: readable  # Document Style
    highlight: tango # Syntax Highlighting
    number_sections: yes # Table Of Content numbering
    toc: yes  # Table Of Content
    toc_depth: 3  # Table Of Content depth
    df_print: paged # Print the data frames in a html table
    code_folding: hide  # Hide the code in the report  
---
  

# Automatic Packages Installing and Loading 
  
```{r setup}
knitr::opts_chunk$set(echo       = TRUE,
                      fig.height = 3.5,
                      fig.width  = 6)

packages <- c("tidyverse", "ggpubr")

## Install packages From CRAN if they are not installed.
for (p in packages) {
  if (p %in% rownames(installed.packages()) == FALSE) {
    install.packages(p, repos = 'http://cran.us.r-project.org')
  }
}
## Load the packages
for (p in packages)
  suppressPackageStartupMessages(
    library(p, quietly = TRUE, character.only = TRUE ) 
  )
```


# Read Data

```{r}
(data_online_scaled <- readRDS(file = "data_online_scaled.RDS")) %>% data.frame()
```


# Some intuition

Number of Unique Values for ID variables: StockCode, InvoiceNo and CustomerID
```{r}
length(unique(data_online_scaled[,'StockCode']))
length(unique(data_online_scaled[,"InvoiceNo"]))
length(unique(data_online_scaled[,"CustomerID"]))
```


# Run the algorithm

```{r}
K <- 3 
set.seed(12)
k_means <- kmeans(x = data_online_scaled, centers = K, nstart = 20, )
str(k_means)
```


# How many transactions per cluster do we have ? 

```{r}
table(k_means$cluster)
```


# Visualization {.tabset}

**On the original scale**
```{r}
data_online_final <- readRDS(file = "data_online_final.RDS")

set.seed(1234)  # Keep same sampling as before
samples_to_keep <- sample(x = 1:nrow(data_online_final), size = 1000)
data_online_final <- data_online_final %>% slice(samples_to_keep)

data_online_final$cluster <- as.factor(k_means$cluster)
```


## ID features {-}

**StockCode**
```{r, warning=FALSE}
# ggplot(data_online_final, aes(x = StockCode, fill = cluster)) +
#   geom_bar() 

ggplot(data_online_final, aes(x = as.numeric(StockCode), fill = cluster, col = cluster)) +
  geom_density(alpha = 0.2, size = 1) +
   # Remove horizontal line in the density plot
  geom_hline(yintercept = 0, colour = "white", size = 1.1)
```

**InvoiceNo**
```{r}
# ggplot(data_online_final, aes(x = InvoiceNo, fill = cluster)) +
#   geom_bar() 

ggplot(data_online_final, aes(x = as.numeric(InvoiceNo), fill = cluster, col = cluster)) +
  geom_density(alpha = 0.2, size = 1)  +
   # Remove horizontal line in the density plot
  geom_hline(yintercept = 0, colour = "white", size = 1.1)
```

**CustomerID**
```{r}
# ggplot(data_online_final, aes(x = as.factor(CustomerID), fill = cluster)) +
#   geom_bar() 

ggplot(data_online_final, aes(x = CustomerID, fill = cluster, col = cluster)) +
  geom_density(alpha = 0.2, size = 1) +
   # Remove horizontal line in the density plot
  geom_hline(yintercept = 0, colour = "white", size = 1.1)
```



## Country {-}

```{r}
ggplot(data_online_final, aes(x = Country, fill = cluster)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
## Without UK 
ggplot(data_online_final %>% filter(Country != "United Kingdom"), aes(x = Country, fill = cluster)) +
  geom_bar() +
  labs(title = "Without United Kingdom") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
ggplot(data_online_final, aes(x = Country, y = Quantity, col = cluster)) +
  geom_jitter() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Price & Quantity {-}

```{r, fig.width = 9}
g_price <- ggplot(data_online_final,
                  aes(x = UnitPrice, fill = cluster, col = cluster)) +
  geom_density(alpha = 0.2, size = 1) +
   # Remove horizontal line in the density plot
  geom_hline(yintercept = 0, colour = "white", size = 1.1)
g_quantity <- ggplot(data_online_final,
                     aes(x = Quantity, fill = cluster, col = cluster)) +
  geom_density(alpha = 0.2, size = 1) +
   # Remove horizontal line in the density plot
  geom_hline(yintercept = 0, colour = "white", size = 1.1)

ggarrange(
  g_price,
  g_quantity,
  common.legend = TRUE,
  legend        = "bottom"
)
```

**Remove "outliers" for better visualization**
```{r, fig.width = 9}
g_price2 <- ggplot( 
  data_online_final %>% filter(UnitPrice < 10),
  aes(x = UnitPrice, fill = cluster, col = cluster)
) +
  geom_density(alpha = 0.2, size = 1) +
   # Remove horizontal line in the density plot
  geom_hline(yintercept = 0, colour = "white", size = 1.1)
g_quantity2 <- ggplot(
  data_online_final %>% filter(Quantity < 150),
  aes(x = Quantity, fill = cluster, col = cluster)
) +
  geom_density(alpha = 0.2, size = 1) + 
  # Remove horizontal line in the density plot
  geom_hline(yintercept = 0, colour = "white", size = 1.1)

ggarrange(
  g_price2,
  g_quantity2,
  common.legend = TRUE,
  legend        = "bottom"
)
```


## Hour of day (7-20) {-}

```{r}
ggplot(data_online_final, aes(hour_of_day, fill = cluster)) +
  geom_bar(position = "fill") 
```

## Day of the Week  {-}

```{r}
ggplot(data_online_final, aes(as.factor(day_of_week), fill = cluster)) +
  geom_bar(position = "fill")
```



