---
title: "Section 1 video 4 : Data Exploration with Online Retail Data Set Sample"
author: "Antoine Pissoort"
date: "February 28, 2018"
output: 
  html_document:  
    theme: readable  # Document Style
    highlight: tango # Syntax Highlightiing
    number_sections: yes # Table Of Content numbering
    toc: yes  # Table Of Content
    toc_depth: 3  # Table Of Content depth
    df_print: paged # Print the data frames in a html table
---

```{r}
knitr::opts_chunk$set(echo = TRUE, fig.height = 1.5)
```


## Automatic Packages Installing and Loading 

```{r}
packages <- c("readr", "tidyverse", "lubridate")

## Install packages From CRAN if they are not installed.
for (p in packages) {
  if (p %in% rownames(installed.packages()) == FALSE) {
    install.packages(p, repos = 'http://cran.us.r-project.org')
  }
}
## Load the packages
for (p in packages)
  suppressPackageStartupMessages(
    library(p, quietly = TRUE, character.only =TRUE ) 
    )
```


# Data Import 

```{r}
data_online <- read_csv("data.csv")
data_online
```

* `InvoiceNo`  : Invoice number. Nominal, a 6-digit integral number uniquely assigned to each transaction. If this code starts with letter 'c', it indicates a cancellation.

* `StockCode`  : Product (item) code. Nominal, a 5-digit integral number uniquely assigned to each distinct product.

* `Description`: Product (item) name. Nominal.

* `Quantity`   : The quantities of each product (item) per transaction. Numeric.

* `InvoiceDate`: Invoice Date and time. Numeric, the day and time when each transaction was generated.

* `UnitPrice`  : Unit price. Numeric, Product price per unit in sterling.

* `CustomerID` : Customer number. Nominal, a 5-digit integral number uniquely assigned to each customer.

* `Country`    : Country name. Nominal, the name of the country where each customer resides. 



# Data pre-processing

 **Missing Values**
```{r}
apply(data_online, 2, function(x) sum(is.na(x)) / length(x) * 100) 
```


 **Duplicated Entries**
```{r}
sum(duplicated(data_online))

data_online %>% filter(duplicated(data_online))
```


**Date time conversion **
```{r}
data_online_time <- data_online %>%
  mutate(
    InvoiceDate = mdy_hm(InvoiceDate),
    day_of_week = wday(InvoiceDate, label = TRUE),
    hour_of_day = hour(InvoiceDate)
  )
```


**Final Dataset**
```{r}
data_online_final <- data_online_time %>%
  mutate(  # Put to missing the negative quantity or price values
    Quantity  = replace(Quantity, Quantity <= 0, NA),
    UnitPrice = replace(UnitPrice, UnitPrice <= 0, NA)
  ) %>%
  filter(complete.cases(.) & !duplicated(.)) %>% 
  mutate(total_price = Quantity * UnitPrice)

# Save the final R dataframe to re-use in the next sections
saveRDS(data_online_final, file = "data_online_final.RDS")
```


# Data exploration 


** Number of distinct invoices, customers and products **
```{r}
data_online_final %>%
  select(InvoiceNo, CustomerID, StockCode) %>%
  apply(., 2, function(x) length(unique(x)))
```



```{r}
# How many different orders per transaction ?
data_online_final %>%
  group_by(InvoiceNo) %>%
  summarise(n = n()) %>%
  mutate(mean = mean(n)) 
```

```{r}
# How many different order per customers ?
data_online_final %>% 
  group_by(CustomerID) %>%
  summarise(n = n()) %>%
  mutate(mean = mean(n)) 
```


```{r}
# How many transactions  per customer ? 
data_online_final %>% 
  group_by(CustomerID) %>% 
  summarise(n = length(unique(InvoiceNo))) %>%
  mutate(mean = mean(n)) %>% 
  arrange(desc(n))
```


```{r}
# How many transactions per countries ? 
data_online_final %>% group_by(Country) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```


**Cancelled orders**

```{r}
sum(substr(data_online$InvoiceNo, 1, 1) == "C")
```

```{r}
data_online %>% 
  group_by(CustomerID) %>%
  summarise(cancelled = sum(substr(InvoiceNo, 1, 1) == "C") ) %>%
  arrange(desc(cancelled))
```


## Some Plots


```{r}
# Homogeneous Theme for ggplot
"theme_piss" <- function(size_p = 18,
                         size_c = 14,
                         size_l = 12,
                         theme = theme_bw(),
                         col = "#33666C",
                         ...) {
  text <-
    function(size, ...)
      element_text(size = size,
                   colour = col,
                   face = "bold",
                   ...)
  theme +
    theme(
      plot.title = text(size_p, hjust = 0.5),
      axis.title = text(size_c),
      legend.title = text(size_l)
    ) +
    theme(
      legend.background = element_rect(colour = "black"),
      legend.key = element_rect(fill = "white"),
      ...
    )
}

# Set the theme for all following plots
theme_set(theme_piss())
```


### Revenue over time 

```{r, fig.height=3.5, fig.width=8}
data_online_final %>%
  mutate(InvoiceDate = as.Date(as.POSIXct(InvoiceDate, "GMT"))) %>% 
  group_by(InvoiceDate) %>%
  summarise(revenue = sum(total_price)) %>%
  ggplot(aes(x = InvoiceDate, y = revenue)) + 
  geom_line() + 
  geom_smooth(method = 'loess', se = FALSE) +
  labs(x = 'Date', y = 'Revenue (£)', title = 'Revenue over time with a smoothed model')
```



### Revenue per days of the week 

```{r, fig.height=3.5, fig.width=8}
data_online_final %>%
  group_by(day_of_week) %>%
  summarise(revenue = sum(total_price)) %>%
  ggplot(aes(x = day_of_week, y = revenue)) + 
  geom_col() + 
  labs(x = 'Day of Week', y = 'Revenue (£)', title = 'Revenue by Day of Week')
```

### Revenue per hours of a day 

```{r, fig.height=3.5, fig.width=8}
data_online_final %>%
  group_by(hour_of_day) %>%
  summarise(revenue = sum(total_price)) %>%
  ggplot(aes(x = hour_of_day, y = revenue)) +
  geom_col() + 
  labs(x = 'Hour Of Day', y = 'Revenue (£)', title = 'Revenue by Hour Of Day')
```



### Distribution of the revenues per Customer 

```{r, fig.height=3.5, fig.width=8}
data_online_final %>%
  group_by(CustomerID) %>%
  summarise(revenue = sum(total_price), 
            transactions = n_distinct(InvoiceNo)) %>%
  mutate(avg_ordr_val = (round((revenue / transactions), 2))) %>% 
  ggplot(aes(x = avg_ordr_val)) + 
  geom_histogram() + 
  scale_x_log10() +
  labs(x = 'Revenue', y = 'Count of Customers', title = 'Histogram of Revenue per customer (Log Scale)')
```


```{r, fig.height=3.5, fig.width=8}
data_online_final %>%
  group_by(CustomerID, Country) %>%
  summarise(revenue = sum(total_price), 
            transactions = n_distinct(InvoiceNo)) %>%
  mutate(avg_ordr_val = (round((revenue / transactions), 2))) %>% 
  filter(Country %in% c("United Kingdom", "Netherlands", "France", "Germany")) %>%
  ggplot(aes(x = avg_ordr_val)) + 
  geom_density(aes(fill = Country, col = Country ), alpha = 0.1, size = 1.2) + 
  scale_x_log10() +
  labs(x = 'Revenue', y = 'Count of Customers', title = 'Histogram of Revenue per customer (Log Scale)')
```

